A Data-Driven Framework for Neural Field Modelling

### 2010-11-02 ltx/first_order_framework/LaTeX_Manuscript.tex
correction to the table
correction to figure referencing

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} In this section, we provide a demonstration of the estimation procedure using the new methods that are described above. The full neural field model was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field can be seen in Figure~\ref{fig:Figure3}(A). Figure~\ref{fig:Figure3}(B) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(C) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis} 
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:Figure4}(A) shows the spatial frequency of the neural field and Figure~\ref{fig:Figure4}(B) shows the spatial frequency of the reconstruct neural field from the basis function decomposition. The cut-off frequency of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing.

The spatial frequency of the observed field is shown in Figure~\ref{fig:Figure5}(A). The cut-off frequency is $\boldsymbol{\nu}_{cy} \approx 0.2$~Hz (-3~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:Figure5}(B). Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be 2.5~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to these criteria.

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(C). The band-limiting effect of the basis function decomposition can be seen when comparing Figure~\ref{fig:Figure4}(A) to Figure~\ref{fig:Figure4}(B).

@@

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for \dean{150} realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range ([-3.29,3.77]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(B) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(A)), but is missing the higher-frequency detail.

@@

\small{\begin{tabular}{lccccccccc}
Experiment&\multicolumn{3}{c}{IV}&\multicolumn{3}{c}{V}&\multicolumn{3}{c}{VI}\\
 Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\theta_0$&$\theta_1$&$\theta_2$&$\theta_0$&$\theta_1$&$\theta_2$\\
\hline\hline
Actual Value&$15$&$0$&$0$&$30$&$0$&$0$&$50$&$0$&$0$\\
Mean Estimate&$35.68$&$-18.16$&$0.81$&$40.43$&$-12.01$&$ 0.35$&$61.46$&$-15.03$&$0.27$\\
Std. Dev.&$16.72$&$11.77$&$0.53$&$11.47$&$7.8$&$0.36$&$11.51$&$7.56$&$0.29$\\
Bias (\%)&$22.97$&$22.7$&$16.2$&$10.43$&$15.01$&$7$&$0.42$&$18.78$&$5.4$\\
\end{tabular}}

>>>>

% Results and Discussion can be combined.
\section{Results}\label{ResultsSection} In this section, we provide a demonstration of the estimation procedure using the new methods that are described above. The full neural field model was used to generate the data for state and parameter estimation. All parameters for the model are given in Table~\ref{tab:Parameters}. Note, the parameters for the spatial and temporal aspects of the model can be considered arbitrary when demonstrating the estimation methodology. This is due to the fact that the temporal aspects scale with the sampling rate and the length of the times series, and the spatial aspects scale with the size of the neural field and the space step used for numerical simulation. When generating the simulated data, we assumed free boundary conditions for the neural field. An example of the simulated field can be seen in Figure~\ref{fig:Figure3}(a). Figure~\ref{fig:Figure3}(b) shows an example of data from the modelled sensors and Figure~\ref{fig:Figure3}(c) shows the power spectral density (PSD) of the simulated time-series. %The PSD shows the typical $1/f^2$ characteristics of intracranial EEG.

@@

\subsection{Spatial Frequency Analysis} 
The spectral analysis was used to confirm that the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{fig:Figure4}(a) shows the spatial frequency of the neural field and Figure~\ref{fig:Figure4}(b) shows the spatial frequency of the reconstruct neural field from the basis function decomposition. The cut-off frequency of the neural field is $\boldsymbol{\nu}_c \approx 0.24$~Hz ($-3$~dB point). From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of 2.08~mm. This confirms that the separation of $\Delta_{y} = 1.5$~mm was sufficient to mitigate problems associated with spatial aliasing.

The spatial frequency of the observed field is shown in Figure~\ref{fig:Figure5}(a). The cut-off frequency is $\boldsymbol{\nu}_{cy} \approx 0.2$~Hz (-3~dB point). The observation cut-off frequency is lower than the field cut-off due to the low-pass filtering effect of the sensors. The spatial frequency response of the sensors can be seen in Figure~\ref{fig:Figure5}(b). Using an oversampling parameter of $\rho_{\phi}=1$, the minimum distance between the field basis functions was calculated to be 2.5~mm to satisfy Shannon's sampling criterion (from equation~\ref{eq:BasisFunctionSeparation}). The distance between basis functions was set to $\Delta_{\phi}=2.5$~mm, which is sufficient to capture the significant dynamics according to these criteria.

Given this basis function separation, the basis function width parameter was set to $\sigma_{\phi}^2=2.5$~mm. This allowed a smooth representation of the field by the basis function decomposition up to a spatial frequency $\boldsymbol{\nu}_{c\phi} \approx 0.12$~Hz, as described by equation~\ref{eq:CutoffFromBasisFuncWidth}. The band-limiting effect is due to the spatial frequency response of the basis functions as seen in Figure~\ref{fig:Figure5}(c). The band-limiting effect of the basis function decomposition can be seen when comparing Figure~\ref{fig:Figure4}(a) to Figure~\ref{fig:Figure4}(b).

@@

The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over iterations) of the root mean square error (MRMSE) over space. The MRMSE of the field estimates over time for \dean{150} realisations is shown in Figure~\ref{fig:Figure9}, with the $95~\%$ confidence interval. At each time point, the error is consistently around 0.5~mV, which is approximately 7.1~\% of the field voltage range ([-3.29,3.77]~mV). The error in the reconstructed field is mostly due to the loss of high spatial frequency information from the basis function decomposition as illustrated in Figure~\ref{fig:Figure10}. The figure shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:Figure10}(b) that the estimated field contains the basic structure of the true field (shown in Figure~\ref{fig:Figure10}(a)), but is missing the higher-frequency detail.

@@

\small{\begin{tabular}{lccccccccc}
Experiment&\multicolumn{3}{c}{IV}&\multicolumn{3}{c}{V}&\multicolumn{3}{c}{VI}\\
 Parameters&$\theta_0$&$\theta_1$&$\theta_2$&$\theta_0$&$\theta_1$&$\theta_2$&$\theta_0$&$\theta_1$&$\theta_2$\\
\hline\hline
Actual Value&$15$&$0$&$0$&$30$&$0$&$0$&$50$&$0$&$0$\\
Mean Estimate&$35.68$&$-18.16$&$0.81$&$40.43$&$-12.01$&$ 0.35$&$61.46$&$-15.03$&$0.27$\\
Std. Dev.&$16.72$&$11.77$&$0.53$&$11.47$&$7.8$&$0.36$&$11.51$&$7.56$&$0.29$\\
Bias (\%)&$137.87$&$-$&$-$&$34.77$&$-$&$-$&$22.92$&$-$&$-$\\
\end{tabular}}
